---
title: "Case study: scoring a CDISC pilot for submission readiness"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Case study: scoring a CDISC pilot for submission readiness}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 7)
```

You have run Pinnacle 21, checked your Define-XML, and looked over the datasets.
The output is a few hundred findings spread across a dozen tools. The question
the study team actually asks is simpler than any single report answers: are we
ready to file?

This is what the Submission Confidence Index (SCI) is for. It takes a table of
evidence and returns one number between 0 and 100, plus a decision band, plus a
breakdown of where the number comes from. Here we walk a realistic CDISC pilot
study through it, using the example evidence that ships with `r4subdata`.

```{r libs, message = FALSE}
library(r4subscore)
library(r4subdata)
```

## The evidence

`evidence_pharma` is a validation run for a CDISC pilot study: 250 findings from
tools like Pinnacle 21, spread across the four readiness pillars.

```{r evidence}
ev <- evidence_pharma
nrow(ev)
table(ev$indicator_domain, ev$result)
```

Each row is one check against one asset, carrying a severity and a result. That
is the whole contract. Any tool that can write this shape feeds the score.

## The score

Three calls take the evidence to a number: indicators, pillars, then the
composite index.

```{r score}
ind <- compute_indicator_scores(ev)
ps  <- compute_pillar_scores(ev)
sci <- compute_sci(ps)

sci
```

The pillar scores show where the study stands on each front before they are
weighted together.

```{r pillars}
ps
```

## Where the score comes from

A single number is only useful if you can open it up. `sci_explain()` reports
each pillar's contribution and, more usefully, ranks the indicators dragging the
score down.

```{r explain}
expl <- sci_explain(ev)
expl$pillar_contributions
head(expl$indicator_contributions, 8)
```

The indicators at the top of that list are where remediation buys the most. A
team can work the list from the top and watch the score move, rather than
guessing which of 250 findings matter.

## Is the score stable?

The default weights (quality 0.35, trace 0.25, risk 0.25, usability 0.15) are a
choice, not a law. Before anyone leans on the number, it is worth asking how much
it depends on that choice. `sci_sensitivity_analysis()` recomputes the SCI across
a grid of alternative weightings.

```{r sensitivity}
grid <- data.frame(
  quality   = c(0.35, 0.50, 0.25, 0.40),
  trace     = c(0.25, 0.20, 0.30, 0.20),
  risk      = c(0.25, 0.20, 0.25, 0.30),
  usability = c(0.15, 0.10, 0.20, 0.10)
)

sci_sensitivity_analysis(ev, weight_grid = grid)
```

If the band holds across that grid, the readiness call is robust to the
weighting. If it flips between bands, the honest reading is that the decision
hangs on a judgment about which pillar matters most, and that judgment should be
made explicitly rather than buried in a default.

## What this answers

The SCI does not replace the underlying checks. It sits on top of them and
answers the one question the checks cannot: taken together, how ready is this
submission, and what would move the needle first. For a study team deciding
whether to file, that is the number that matters.
