---
title: "Submission Confidence Index (SCI) Scoring"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Submission Confidence Index (SCI) Scoring}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

The `r4subscore` package computes the **Submission Confidence Index (SCI)**,
a composite 0-100 score summarising submission readiness across four pillars:
quality, trace, risk, and usability.

```{r load}
library(r4subscore)
```

## Default configuration

`sci_config_default()` returns the pillar weights and decision-band thresholds:

```{r config}
cfg <- sci_config_default()
cfg$pillar_weights
cfg$bands
```

The default weights are: quality 35%, trace 25%, risk 25%, usability 15%.

## Decision bands

| Band | SCI range | Meaning |
|------|-----------|---------|
| `ready` | 85-100 | Submission ready |
| `minor_gaps` | 70-84 | Minor gaps to address |
| `conditional` | 50-69 | Conditional — significant work needed |
| `high_risk` | 0-49 | High risk — submission not advised |

## Building example evidence

```{r evidence}
ev <- data.frame(
  run_id           = "run-001",
  study_id         = "STUDY01",
  asset_type       = "dataset",
  asset_id         = rep(c("ADSL", "ADAE"), each = 6),
  source_name      = "r4subcore",
  source_version   = "0.1.2",
  indicator_id     = paste0("X-00", 1:12),
  indicator_name   = paste0("Indicator ", 1:12),
  indicator_domain = rep(c("quality", "trace", "risk", "usability",
                           "quality", "trace"), 2),
  severity         = "info",
  result           = c("pass", "pass", "pass", "pass", "pass", "pass",
                       "warn", "pass", "fail", "pass", "pass", "warn"),
  metric_value     = c(1, 1, 1, 1, 1, 1, 0.8, 1, 0, 1, 1, 0.9),
  metric_unit      = "score",
  message          = paste0("Indicator ", 1:12, " result"),
  location         = rep(c("ADSL", "ADAE"), each = 6),
  evidence_payload = "{}",
  created_at       = Sys.time(),
  stringsAsFactors = FALSE
)
```

## Computing pillar scores

`compute_pillar_scores()` aggregates the mean metric value per domain:

```{r pillar-scores}
ps <- compute_pillar_scores(ev)
ps
```

## Computing the SCI

```{r sci}
result <- compute_sci(ps)
result$SCI
result$band
print(result)
```

## Detailed explanation

`sci_explain()` returns a breakdown showing each pillar's contribution to
the final SCI and any score loss:

```{r explain}
expl <- sci_explain(ev)
expl$pillar_contributions
```

## Sensitivity analysis

`sci_sensitivity_analysis()` evaluates the SCI across a grid of alternative
pillar weights, useful for understanding how robust the score is to weighting
choices:

```{r sensitivity}
grid <- data.frame(
  quality    = c(0.35, 0.50, 0.25),
  trace      = c(0.25, 0.20, 0.30),
  risk       = c(0.25, 0.20, 0.25),
  usability  = c(0.15, 0.10, 0.20)
)

sa <- sci_sensitivity_analysis(ev, weight_grid = grid)
sa
```

## Custom weights

Pass a custom config to favour particular pillars:

```{r custom-weights}
cfg_custom <- sci_config_default(
  pillar_weights = c(quality = 0.50, trace = 0.20,
                     risk = 0.20, usability = 0.10)
)
ps2   <- compute_pillar_scores(ev, config = cfg_custom)
sci2  <- compute_sci(ps2, config = cfg_custom)
sci2$SCI
sci2$band
```

## Confidence intervals

A single SCI number hides how much of the score is measurement noise. An SCI of
72 could be a solid `conditional` result or a borderline `minor_gaps` one that
happened to round down. `sci_confidence_interval()` resamples the evidence and
reports where the SCI is likely to sit:

```{r ci}
ci <- sci_confidence_interval(ev, n_boot = 500, seed = 1)
ci
```

`format_sci_ci()` gives the compact form for a report or a slide:

```{r ci-format}
format_sci_ci(ci)
```

Read the interval, not just the point estimate. When the interval sits entirely
inside one decision band, the band is a firm call. When it straddles two bands,
as flagged in the printout, the readiness decision is sensitive to which checks
were run, and the honest answer is that more evidence is needed before the band
can be trusted. The default scheme resamples indicator scores within each
pillar; pass `by = "evidence"` to resample the raw evidence rows instead.

## Critical findings cap the band

The SCI is a weighted average, so a strong pillar can lift the score even when
another pillar hides a serious problem. That is the wrong behavior for a
readiness call: a single critical failure should not be averaged away. The
scoring config carries a non-compensatory gate that caps the band when an open
critical finding is present. The pipeline applies it automatically, counting
critical findings (severity `critical`, result `fail`) from the evidence, but it
is easiest to see with the count passed directly:

```{r gate}
strong <- data.frame(
  pillar       = c("quality", "trace", "risk", "usability"),
  pillar_score = c(0.95, 0.92, 0.90, 0.93),
  weight       = c(0.35, 0.25, 0.25, 0.15)
)

compute_sci(strong, n_critical = 0)$band

capped <- compute_sci(strong, n_critical = 1)
capped$band
capped$gate_reason
```

The numeric SCI is identical in both cases; only the band moves, so the score
still tells you how much work remains. A sponsor that wants a different policy
can move the cap or turn the gate off through `sci_config_default(gate = ...)`.

## Tracking readiness over time

A single score answers "are we ready today". The question a team actually lives
with is "are we on track to file". Snapshots record the score at each dry run so
successive runs can be compared. Here two runs on the same study, with one
traceability failure fixed between them:

```{r snapshots}
ctx <- r4subcore::r4sub_run_context("STUDY1", "DEV")

run_evidence <- function(trace_result) {
  r4subcore::as_evidence(
    data.frame(
      asset_type = "dataset", asset_id = "ADSL", source_name = "pinnacle21",
      indicator_id = c("Q-LABEL", "T-TRACE", "R-RPN"),
      indicator_name = c("Labels present", "ADaM traces to SDTM", "Risk in range"),
      indicator_domain = c("quality", "trace", "risk"),
      severity = c("low", "high", "medium"),
      result = c("pass", trace_result, "pass"),
      stringsAsFactors = FALSE
    ),
    ctx = ctx
  )
}

week1 <- sci_snapshot(run_evidence("fail"), "DB lock minus 8 weeks")
week2 <- sci_snapshot(run_evidence("pass"), "DB lock minus 4 weeks")

week1
week2
```

`sci_diff()` reports what moved and why:

```{r diff}
d <- sci_diff(week1, week2)
d
d$indicators[, c("indicator_id", "status", "delta")]
```

`sci_snapshot_history()` binds a series into one row per run, ready for a trend
plot:

```{r history}
sci_snapshot_history(list(week1, week2))
```
