---
title: "Validating the Submission Confidence Index"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Validating the Submission Confidence Index}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 6, fig.height = 4)
```

A composite score is only useful if it behaves the way its users expect. This
article documents how the Submission Confidence Index (SCI) was checked against
a set of validity criteria, and it is honest about the one criterion that cannot
be checked with public data.

```{r load}
library(r4subscore)
```

## What can and cannot be validated

The ideal test of a readiness score would compare it against real regulatory
outcomes: score a large set of historical submissions, then check whether low
scores predicted refuse-to-file letters, information requests, or review delays.
That study is not possible here. Regulators do not publish per-submission
outcomes linked to dataset-level quality metrics, so there is no public
ground-truth label to fit or test against. Claiming predictive validity without
that data would be dishonest.

What can be established, and what this article covers, is the validity that does
not need outcome labels:

- **Construct validity**: the score falls as evidence gets worse, and never
  rises. This is checked with a degradation curve.
- **Discriminant validity**: a clean package and a broken package land in clearly
  different decision bands.
- **Concurrent validity**: the score agrees with an independent count of
  conformance findings, the kind of figure a reviewer reads off a validation
  report.
- **Robustness**: the score and its ranking are stable under reasonable changes
  to the pillar weights.

The final section is explicit about what a predictive study would need and points
to a worked example on real public data.

## A worked evidence table

The checks below use a single evidence table with a spread of domains,
severities, and results. Everything is built from base data frames so this
article is fully reproducible.

```{r build-evidence}
make_evidence <- function(domain, severity, result, id_prefix = "IND") {
  ctx <- suppressMessages(
    r4subcore::r4sub_run_context(study_id = "VALDEMO", environment = "DEV")
  )
  df <- data.frame(
    asset_type       = "dataset",
    asset_id         = "ADSL",
    source_name      = "demo",
    indicator_id     = paste0(id_prefix, "-", seq_along(result)),
    indicator_name   = paste("check", seq_along(result)),
    indicator_domain = domain,
    severity         = severity,
    result           = result,
    stringsAsFactors = FALSE
  )
  suppressMessages(r4subcore::as_evidence(df, ctx = ctx))
}

domains    <- rep(c("quality", "trace", "risk", "usability"), each = 5)
severities <- rep(c("info", "low", "medium", "high", "critical"), times = 4)
results    <- rep("pass", 20)

evidence <- make_evidence(domains, severities, results)
compute_sci(compute_pillar_scores(evidence))$SCI
```

## Construct validity: the score responds to degradation

`sci_degradation_curve()` turns a growing number of passing checks into failures
and records the SCI at each step. Checks are degraded in order of their
contribution to the baseline score, so the curve shows the steepest descent the
evidence allows. A valid index must never improve when a check is broken.

```{r degradation}
curve <- sci_degradation_curve(evidence, n_steps = 10)
curve
```

`sci_monotone_check()` confirms the curve never rises. A rising step would be a
defect in the scoring model.

```{r monotone}
chk <- sci_monotone_check(curve)
chk$monotone
chk$max_increase
```

```{r degradation-plot, echo = FALSE}
plot(
  curve$frac_degraded, curve$SCI,
  type = "b", pch = 19,
  xlab = "Fraction of passing checks turned into failures",
  ylab = "SCI",
  main = "SCI response to degradation",
  ylim = c(0, 100)
)
abline(h = c(85, 70, 50), lty = 3, col = "grey60")
```

The dotted lines mark the band thresholds. As failures accumulate the score
crosses from one band into the next, in the expected direction.

## Discriminant validity: clean and broken packages separate

Two packages at the extremes should land in clearly different bands: a clean
package where every check passes, and a broken package with failures
concentrated in the more severe checks.

```{r known-groups}
clean  <- make_evidence(domains, severities, rep("pass", 20), "CLEAN")

broken_results <- rep("pass", 20)
broken_results[severities %in% c("medium", "high", "critical")] <- "fail"
broken <- make_evidence(domains, severities, broken_results, "BROKEN")

clean_sci  <- compute_sci(compute_pillar_scores(clean))
broken_sci <- compute_sci(compute_pillar_scores(broken))

data.frame(
  package = c("clean", "broken"),
  SCI     = c(clean_sci$SCI, broken_sci$SCI),
  band    = c(clean_sci$band, broken_sci$band)
)
```

The clean and broken packages fall in different bands, with a wide gap between
their scores. The clean package does not reach the top band here because its
absolute score is capped by the mix of severities it contains. That cap is a
calibration finding in its own right, examined below. The point of this check is
the separation and ordering, which hold: the clean package scores well above the
broken one.

## Concurrent validity: agreement with an independent count

`conformance_findings()` summarises an evidence table into a single
severity-weighted count of unresolved findings. It does not use the SCI pillar
weights, so it is an independent reference. Across a gradient of packages with a
rising share of failures, a higher SCI should line up with fewer findings.

```{r concurrent}
fractions <- seq(0, 0.9, by = 0.1)

gradient <- lapply(fractions, function(f) {
  res <- rep("pass", 20)
  n_fail <- round(f * 20)
  if (n_fail > 0) res[seq_len(n_fail)] <- "fail"
  ev <- make_evidence(domains, severities, res)
  data.frame(
    fraction_failed = f,
    SCI             = compute_sci(compute_pillar_scores(ev))$SCI,
    findings        = conformance_findings(ev)
  )
})
gradient <- do.call(rbind, gradient)
gradient
```

```{r concurrent-cor}
# Spearman rank correlation between the SCI and the independent finding count.
cor(gradient$SCI, gradient$findings, method = "spearman")
```

The correlation is strongly negative: as the independent finding count rises, the
SCI falls. The two measures agree on ordering while remaining distinct in how
they treat severity, so the SCI is not simply a restatement of the finding count.

## Robustness: stability under alternative weights

`sci_sensitivity_analysis()` recomputes the SCI across a grid of weight
scenarios. A score that flips bands under small, defensible weight changes would
be fragile.

```{r robustness}
grid <- data.frame(
  quality   = c(0.35, 0.40, 0.25, 0.30),
  trace     = c(0.25, 0.20, 0.30, 0.25),
  risk      = c(0.25, 0.25, 0.25, 0.30),
  usability = c(0.15, 0.15, 0.20, 0.15)
)

partial <- make_evidence(domains, severities,
                         c(rep("fail", 6), rep("pass", 14)))
sa <- sci_sensitivity_analysis(partial, weight_grid = grid)
sa

# Spread of the SCI across all weight scenarios.
diff(range(sa$SCI))
```

A small spread across scenarios, and a stable band, mean the score does not hinge
on one particular choice of weights.

## A calibration finding

The checks above all concern how the score behaves as evidence changes. Running
them surfaced a separate point about the score's absolute level that is worth
recording plainly.

The scoring model gives a passing check a contribution of `1 - severity_weight`.
Severity therefore discounts the credit a check earns for passing, rather than
weighting the penalty it incurs for failing. The effect is easiest to see with a
package where every check passes, scored at each severity level in turn.

```{r calibration}
by_severity <- lapply(
  c("info", "low", "medium", "high", "critical"),
  function(s) {
    ev <- make_evidence(
      domain   = c("quality", "trace", "risk", "usability"),
      severity = rep(s, 4),
      result   = rep("pass", 4)
    )
    r <- compute_sci(compute_pillar_scores(ev))
    data.frame(severity = s, SCI = r$SCI, band = r$band)
  }
)
do.call(rbind, by_severity)
```

Every check passes in all five packages, yet the score falls from 100 to 0 as
severity rises, and an all-passing high-severity package is labelled `high_risk`.
That is the opposite of the intended reading. It happens because severity
discounts passing credit instead of amplifying failure penalties.

A more standard formulation would score each check as
`1 - (1 - result_score) * severity_weight`, so a pass always scores 1 regardless
of severity, an info failure is harmless, and a critical failure scores 0. That
change is left for a future release: it alters every existing score, so it should
be made deliberately and, ideally, calibrated against the outcome data described
next rather than adopted on intuition alone. The finding is recorded here so the
current behaviour is documented and testable.

## The limit of this validation

Everything above is construct, discriminant, concurrent, and robustness validity.
None of it is predictive validity against regulatory decisions, because that data
is not public.

## What a predictive study would need

A genuine outcome study would need historical submissions paired with their
review outcomes: information requests, review-clock extensions, or filing
decisions, each linked to the evidence available at submission time. That
requires a sponsor consortium willing to contribute past submissions and their
outcomes under a shared agreement. Until such a dataset exists, the honest claim
for the SCI is the one this article supports: it responds correctly to evidence
quality, separates good packages from bad ones, agrees with an independent
finding count, and is stable to reasonable weighting choices.

For a worked example on real public data, see the CDISC pilot case study in the
`r4subpharma` package, which scores an ADaM package built by `admiral` against
its `metacore` specification.
